

>***Analysis of how chemical properties influence the quality of red wines 
by Meixian Chen***


> **Introduction**: This report uses R to quantitatively analyze how chemical 
properties influence the quality of red wines. The tidy wine data consist of 
1599 red wines with 11 variables on the chemical properties of the wine. 
Each wine is rated by at least 3 wine experts, providing a rating between 
0 (very bad) and 10 (very excellent).

> **Overview of the wine dataset**


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

# install.packages("..."") if necessary 

library(ggplot2)
library(GGally)
library(reshape2)
library(gridExtra)
library(grid)
library(corrplot)
library(scatterplot3d)
library(RColorBrewer)
```

> Here is part of the dataset, contains 1599 observations of 11 chemical 
properties 
and quality ratings of wines.

```{r echo=FALSE, Load_the_Data}
# Load the Data
red = read.csv("wineQualityReds.csv",header = T)
red$X <- NULL
str(red)


```

> Summary of statistics of each variables

```{r,echo=FALSE}
summary(red)
```


# Univariate Plots Section

> Visual distribution of each variables

```{r, echo=FALSE}
q1 = ggplot(aes(x= quality), data = red)+
  geom_histogram(binwidth = 0.5, color='gray')+
  stat_bin(binwidth = 1,aes(y=..count.., label=..count..), geom="text", vjust=-.5) +
  scale_x_continuous(breaks=c(3,4,5,6,7,8))+
  ggtitle("Distribution of the ranking of red wine qualities")
q1
```

> The distribution of wine quality is close to normal distribution.
Most wines are rated as average (5 or 6), there are a few bad wines (less than 
5) and good wines (more than 7). 

> We continue to inspect the distribution of each of the 11 variables. 

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x   = fixed.acidity  ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = fixed.acidity ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> The distribution of fixed.acidity is close to normal distribution. 
The fixed acidity values above 12.5% are considered as outliers. 


```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = volatile.acidity  ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = volatile.acidity ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> I remove the outlier points which has a close to 1.6% volatile acidity from
the dataframe in the further analysis.

```{r, echo=FALSE}
red = red[red$volatile.acidity<1.4,]
```

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = citric.acid  ) ) + 
                   geom_histogram(binwidth = 0.08, color='gray' ),
              ggplot(red, aes( x = 1, y = citric.acid ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> Different from the above two variables, the distribution of citric.acid 
has a peak value in 0, and the numbers decrease with the citric.acid values. 
I remove the outliers which value is 1.00. 

```{r,echo=FALSE}
red = red[red$citric.acid<0.9,]
```

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = residual.sugar  ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = residual.sugar ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> Most of the residual.sugar level are between [0,4]. This distribution has 
a long and small tail.  

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = chlorides  ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = chlorides ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> The distribution of chlorides is very similar to residual.sugar,
and both of them have a long tail of outliers. Are
the outliers in the two plot from the same observation of the data?  

> The next two boxplots are plotting chlorides variable on the data set 
which the residual.sugar level is less than 3, and residual.sugar on the data
set which the chlorides level is less than 1.2. 

```{r}
grid.arrange( 
              ggplot(red[red$residual.sugar<3,], aes( x = 1, y = chlorides ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red') ,
              ggplot(red[red$chlorides<1.2,], aes( x = 1, y = residual.sugar ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ),
            ncol=2)
```

> If the outliers are from some common points, we expect to see the 
change from the original plots. Since there is no obvious change on the new 
plots, the outliers are from different points. Moreover, the correlation value 
cor(red$residual.sugar,red$chlorides) = 0.059 is rather low. 
Similar distributions of two variables do not means
there is a high correlationship between them. 

```{r}
cor(red$residual.sugar,red$chlorides)
```


```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = free.sulfur.dioxide  ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = free.sulfur.dioxide ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = total.sulfur.dioxide  ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = total.sulfur.dioxide ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> I remove outliers with values >200 in the further analysis. 

```{r,echo=FALSE}
red = red[red$total.sulfur.dioxide<200,]
```

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = density ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = density ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> The distribution of red wine density are in a small range between 0.99 g/ml
to 1.01 g/ml (pure water is 1.00 g/ml). It matches what I expected. 

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = pH ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = pH ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> Wine has pH between 2.9 to 4. 

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = sulphates ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = sulphates ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```

> I remove outliers with sulphates values >1.5. 

```{r,echo=FALSE}
red = red[red$sulphates < 1.5,]
```

```{r, echo=FALSE,message=FALSE, warning=FALSE}
grid.arrange( ggplot(red, aes( x = alcohol ) ) + 
                   geom_histogram(color='gray' ),
              ggplot(red, aes( x = 1, y = alcohol ) ) + 
               geom_jitter(alpha = 0.1 ) +
               geom_boxplot(alpha = 0.2, color = 'red' ) ,
            ncol=2)
```


> The alcohol level of red wines is almostly between 8~14 degrees. 

# Univariate Analysis

> The distributions of most variables are close to normal distribution:
the majority of individuals are in the middle slots and fewer are in the 
low/ high slots. 



> We are mainly interested on which chemical properties affect the wine 
quality, and also how much they relate to each other. 
Here is matrix presenting the correlation values among the variables. 
Red color dots shows negative correlation and blue are positive correlation.
The darkness and the size of the dots shows how strong is the correlation. 

```{r,echo=FALSE}
corMatrix =cor(red,method='pearson')
corrplot(corMatrix, type="lower")
```


> From the correlation matrix, quality is mainly related to alcohol and 
volatile.acidity levels, then to citric.acid and sulphates levels.  




# Bivariate Plots and Analysis Section

> We plot the quality variable with the chemical properties which highly 
related to it. 

```{r echo=FALSE, Bivariate_Plots_1}

#factorize quality variable and use it as a parameterize the color 
#in the plotting
red$quality_factor = factor(red$quality)

p_alcohol_quality = ggplot(red, 
       aes(y=alcohol,x=quality,color=quality_factor))+
  geom_jitter()+
  geom_boxplot(aes(group=(cut_width(quality,1))),alpha = 0.5)+
  ylab("Alcohol levels in red wine[%]")+
  coord_flip()+
  scale_color_brewer(type = 'div',palette = 4,guide = FALSE) +
  theme_update(plot.title = element_text(hjust = 0.5))+
  ggtitle("Alcohol levels in red wines VS. quality")
  
p_alcohol_quality
```

>
> By plotting the alcohol and quality, we can see better quality red wines tends 
to have higher alcohol levels. For average and good wines (rated more than 4), 
the mean of alcohol levels increases with the ranking categories. 
 


```{r echo=FALSE, Bivariate_Plots_2}
ggplot(red, 
       aes(y=volatile.acidity,x=quality,color=quality_factor))+
  geom_jitter()+
  geom_boxplot(aes(group=(cut_width(quality,1))),alpha = 0.5)+
  ylab("Volatile acidity level in red wine[g/dm^3]")+
  coord_flip()+
  scale_color_brewer(type = 'div',palette = 4,guide = FALSE) +
  ggtitle("Volatile acidity level in red wines VS. quality")

```

>
> Good wine tends to have low volatile acidity. 

> The third chemical property we try to plot is citric.acid. It is the 
third/fourth variable that relates to quality, and it is also strongly related 
to volatile.acidity. We would first provide bivariate plot of citric.acid and 
quality, and in the next section, further investigate the multi-variable 
relationship among quality and citric.acid and volatile.acidity. 


```{r echo=FALSE, Bivariate_Plots_3}

ggplot(red, 
       aes(y=citric.acid,x=quality,color=quality_factor))+
  geom_jitter()+
  geom_boxplot(aes(group=(cut_width(quality,1))),alpha = 0.5)+
  coord_flip()+
  ylab("Citric acid level in red wine[g/dm^3]")+
  scale_color_brewer(type = 'div',palette = 4,guide = FALSE) +
  ggtitle("Citric acid level in red wines VS. quality")

```

> Good wine tends to have higher citric acid. 


# Multivariate Plots and Analysis Section



```{r echo=FALSE, Multivariate_Plots_1}

ggplot(red, 
       aes(x=alcohol,y=volatile.acidity,color=quality_factor))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE,size=1)+
  scale_color_brewer(type = 'div',palette = 4) +
  xlab("Alcohol level in red wine[%]")+
  ylab("Volatile acidity level in red wine[g/dm^3]")+
  ggtitle("Alcohol and volatile acidity level in red wines VS. quality")
```

>
> The first multivariable plot is a point plot, using the two most strongest 
properties, alcohol and volatile acidity as axis, and quality as color to 
highlight the distribution of different categories of wines. The majority of 
good wines lie in the zone of higher alcohol value and lower volatile acidity 
value. While the bad wines are the opposite case. 


```{r echo=FALSE, warning=FALSE, message=FALSE, Multivariate_Plots_3}
q_multi_3= ggplot(red, 
       aes(x=citric.acid,y=volatile.acidity,color=quality_factor))+
  geom_point(alpha = 0.7)+
  ylim(c(0.2,1))+
  xlim(c(0,0.8))+
  geom_smooth(method = "lm", se = FALSE,size=1)+
  scale_color_brewer(type = 'div',palette = 4) +
  xlab("Citric acid level in red wine[g/dm^3]")+
  ylab("Volatile acidity level in red wine[g/dm^3]")+
  ggtitle("citric acidity and volatile acidity level in red wines VS. quality")

q_multi_3
```


>
> The second multivariable plot uses 
citric acid and volatile acidity as axis, and quality as color to highlight the 
distribution of different categories of wines. The majority of good wines lie 
in the zone of higher citric acid value and higher volatile acidity value. 
While the bad wines are the opposite case. More interesting, wine with higher 
citric acid level tends to have also higher volatile acidity level. 

```{r echo=FALSE, Multivariate_Plots_3d}


red$quality_color[red$quality==3] = 'darkgoldenrod'
red$quality_color[red$quality==4] = 'navajowhite2'
red$quality_color[red$quality==7] = 'plum3'
red$quality_color[red$quality==8] = 'darkorchid4'


with(subset(red, quality<4 | quality>7),{
  wine3d <-scatterplot3d(volatile.acidity,citric.acid,alcohol,
                color=quality_color,
                type='h',
                pch=19, 
                main="3D plot of good and bad wine vs. 
                volatile.acidity,citric.acid and alcohol ")
  wine3d.coords <- wine3d$xyz.convert(volatile.acidity,citric.acid,alcohol)
  legend("topright", inset=.05,      # location and inset
    bty="n", cex=.8,              # suppress legend box, shrink text 50%
    title="Quality of wine",
    c("very good (8)", "very bad (3)"), fill=c("darkgoldenrod", "darkorchid4"))
}
)

```

>
>The last plot we present is a 3D figure. We select only the very good (rating 
maximal 8) and very bad wine (rating minimal 4) from the dataset in order to 
have a clear image. The interesting fact we find is that, good wines usually
have a lower volatile acidity, higher citric acid and higher alcohol level,
while bad wines are the opposite. 



# Final Plots and Summary

> To summary this report, we select three presentive plots, an one-variable
plot showing the distribution of wine quality, a bi-variable plot showing how
a chemical property influences the wine quality, and a multi-variable plot
showing how chemical properties influences the wine quality and also how they 
relate to each other. 

### Plot One
```{r echo=FALSE, Plot_One}
q1
  
```

>
> This first plot gives an overview of wine quality distribution. Wine
quality distribution is close to normal distribution, and it is as we expect:
the majority of wines on the market are average, very good and very bad wines
are few. 


### Plot Two
```{r echo=FALSE, Plot_Two}
p_alcohol_quality
```

>
> The second plot shows how one of the most strongly influencing chemical 
property,alcohol, affects the wine quality. Good wines tend to have a high 
alcohol level in general. 


### Plot Three
```{r echo=FALSE,warning=FALSE, message=FALSE, Plot_Three}
q_multi_3
```

>
> The third plot shows how citric acid and volatile acidity influence the wine
quality. Moreover these two variable is also related to each other. 
The majority of good wines lie 
in the zone of higher citric acid value and higher volatile acidity value. 
While the bad wines are the opposite case. More interesting, wine with higher 
citric acid level tends to have also higher volatile acidity level.


# Reflection

> The most different part when starting the analysis was to decide starting from
which variable. I wasted some time at the beginner while trying to plot 
insightful figures. Generating a correlation matrix among variable, or using
ggpair library to analysis a subset of the dataset in advance of 
bi/multi-variable plotting is very useful. The analysis went more smooths by 
picking up the right combination of variables. Besides, I was also struggled 
on generating clear figures to show the patterns of the data, especially 
on the multi-variable plotting.
By adopting the suggestion of draw linear regression
line of different categories of wine qualities, it is easier to discover 
the patterns. 


> Wine tasting is a personal thing. Some people prefer some kinds of wines 
while the others have different opinions. One future work on the dataset 
analysis could be, storing separately the rating from different experts, and 
then to identify what is the common chemical properties of good wine for each 
individual. Thus, we can recommend good wine based on similar personal 
preference, rather than average opinions of some experts. 

